Affective Uplift During Video Game Play: A Naturalistic Case Study

### 2023-10-27 ms.Rmd
Adjust figures

dir.create("models", FALSE)
```

@@

```{r}
#| label: fig-session-durations
#| include: false
#| fig-height: 3
#| fig.cap: Summaries of session durations. X-axes are truncated at 10 hours.

@@

```{r}
#| label: fig-descriptives
#| fig-height: 2.2
#| fig-width: 7
#| include: false
#| fig.cap: 'A. Histogram of session durations (note log10 x-axis). B. Summary of how many sessions each participant completed. C. Histogram of all mood ratings.'

@@

```{r}
#| label: fig-data
#| include: true
#| fig-width: 8
#| fig-height: 5
#| fig.env: "figure*"
#| fig.cap: "A. Scatterplots of three participants' (rows) mood responses (pre-play: red; during play: blue) over eight sessions' (columns) durations. B. Histograms of session-mean (C. person-mean) moods before (top) and during (bottom) play sessions. D. Differences in session-mean (E. player-mean) mood differences (during session - pre-play). F. Scatterplot of person-mean mood reports at the beginning (x-axis) and during gameplay sessions (y-axis). Identity line is shown in green, and an exploratory GAM regression line is shown in blue."

@@

p_mood_example <- dat_example_sessions |>
  filter(session <= 8) |>
  ggplot() +
  aes(hours, mood, col = post) +
  scale_color_brewer(
    "Pre-session measure",
    palette = "Set1",
    aesthetics = c("color", "fill")
  ) +
  scale_x_continuous(
    "Session duration (hours)"
  ) +
  scale_y_continuous(
    "Mood",
    limits = c(0, 1),
  ) +
  geom_point(size = 1.5, alpha = 1) +
  facet_grid(
    rows = vars(Person),
    cols = vars(Session),
    scales = "fixed"
  ) +
  theme(
    legend.position = "none",
    strip.text.y = element_blank()
  )

# Plot: Raw mood histograms at pre and during play
p_mood_prepost_raw <- dat |>
  mutate(post = factor(post, labels = c("Pre-play", "During play"))) |>
  ggplot(aes(mood, fill = post)) +
  scale_color_brewer(
    "Pre-session measure",
    palette = "Set1",
    aesthetics = c("color", "fill")
  ) +
  geom_histogram(bins = 30, col = "white") +
  scale_y_continuous(
    "Observations",
    expand = expansion(c(0.001, 0.1)),
    breaks = scales::extended_breaks(5)
  ) +
  scale_x_continuous(
    "Mood",
    expand = expansion(c(0.001))
  ) +
  facet_wrap("post", ncol = 1, scales = "free_y") +
  theme(legend.position = "none")

# Plot: Person-session-mean mood histograms at pre and during play
p_mood_prepost_sessions <- p_mood_prepost_raw %+%
  summarise(
    p_mood_prepost_raw$data,
    mood = mean(mood, na.rm = TRUE), .by = c(post, pid, ps)
  ) +
  scale_y_continuous(
    "Sessions",
    expand = expansion(c(0.001, 0.1)),
    breaks = scales::extended_breaks(5)
  )

# Plot: Person-mean mood histograms at pre and during play
p_mood_prepost_players <- p_mood_prepost_sessions %+%
  summarise(
    p_mood_prepost_raw$data,
    mood = mean(mood, na.rm = TRUE), .by = c(post, pid)
  ) +
  scale_y_continuous(
    "Players",
    expand = expansion(c(0.001, 0.1)),
    breaks = scales::extended_breaks(5)
  )

# Plot: Difference histogram (sessions)
p_mood_difference_sessions <- p_mood_prepost_sessions$data |>
  pivot_wider(names_from = post, values_from = mood) |>
  mutate(Difference = `During play` - `Pre-play`) |>
  drop_na(Difference) |>
  ggplot(aes(Difference)) +
  geom_histogram(bins = 50, col = "white") +
  geom_vline(xintercept = 0, linewidth = .5, col = "#2ca25f") +
  scale_y_continuous(
    "Sessions",
    expand = expansion(c(0.001, 0.1))
  ) +
  scale_x_continuous(
    expand = expansion(c(0.001))
  ) +
  coord_cartesian(xlim = c(-.4, .6))

# Plot: Difference histogram (players)
p_mood_difference_players <- p_mood_prepost_players$data |>
  pivot_wider(names_from = post, values_from = mood) |>
  mutate(Difference = `During play` - `Pre-play`) |>
  drop_na(Difference) |>
  ggplot(aes(Difference)) +
  geom_histogram(bins = 50, col = "white") +
  geom_vline(xintercept = 0, linewidth = .5, col = "#2ca25f") +
  scale_y_continuous(
    "Players",
    expand = expansion(c(0.001, 0.1))
  ) +
  scale_x_continuous(
    expand = expansion(c(0.001))
  ) +
  coord_cartesian(xlim = c(-.4, .6))

@@

```{r}
#| label: fig-ct1
#| include: true
#| fig.cap: "Estimated (changes) in mood as a function of session duration. Top: Average mood during a gaming session. Bottom: Change in mood during a session compared to mood at the beginning of a session. Gray ribbons indicate 95\\% confidence bands. We truncated the x-axis at three hours for this figure."
#| fig-height: 3
#| fig-width: 5

@@

```{r}
#| label: fig-ct2
#| fig.env: "figure*"
#| include: true
#| fig.cap: "Estimated (changes) in mood as a function of session duration and pre-play mood. Top. Mood for the average player during a gaming session with pre-play mood at 5th, 25th, 50th, and 75th percentiles (columns). Ribbons indicate 95\\% confidence. Bottom. Same as above but with change in mood on the y-axis."
#| fig-height: 3
#| fig-width: 8

>>>>

dir.create("models", FALSE)

knitr::opts_chunk$set(
  fig.width = 8,
  fig.asp = 0.618
)
```

@@

```{r}
#| label: fig-session-durations
#| include: false
#| fig.asp: 0.5
#| fig.cap: Summaries of session durations. X-axes are truncated at 10 hours.

@@

```{r}
#| label: fig-descriptives
#| fig-height: 2.2
#| include: false
#| fig.cap: 'A. Histogram of session durations (note log10 x-axis). B. Summary of how many sessions each participant completed. C. Histogram of all mood ratings.'

@@

```{r}
#| label: fig-data
#| include: true
#| fig.env: "figure*"
#| fig.cap: "A. Scatterplots of three participants' (rows) mood responses (pre-play: red; during play: blue) over eight sessions' (columns) durations. B. Histograms of session-mean (C. person-mean) moods before (top) and during (bottom) play sessions. D. Differences in session-mean (E. player-mean) mood differences (during session - pre-play). F. Scatterplot of person-mean mood reports at the beginning (x-axis) and during gameplay sessions (y-axis). Identity line is shown in green, and an exploratory GAM regression line is shown in blue."

@@

p_mood_example <- dat_example_sessions |>
  filter(session <= 8) |>
  ggplot() +
  aes(hours, mood, col = post) +
  scale_color_brewer(
    "Pre-session measure",
    palette = "Set1",
    aesthetics = c("color", "fill")
  ) +
  scale_x_continuous(
    "Session duration (hours)"
  ) +
  scale_y_continuous(
    "Mood",
    limits = c(0, 1),
    breaks = c(0, 0.25, 0.5, 0.75, 1.0),
    labels = c("0", "", "0.5", "", "1")
  ) +
  geom_point(size = 1.5, alpha = 1) +
  facet_grid(
    rows = vars(Person),
    cols = vars(Session),
    scales = "fixed"
  ) +
  theme(
    legend.position = "none",
    strip.text.y = element_blank()
  )

# Plot: Raw mood histograms at pre and during play
p_mood_prepost_raw <- dat |>
  mutate(post = factor(post, labels = c("Pre-play", "During play"))) |>
  ggplot(aes(mood, fill = post)) +
  scale_color_brewer(
    "Pre-session measure",
    palette = "Set1",
    aesthetics = c("color", "fill")
  ) +
  geom_histogram(bins = 30, col = "white") +
  scale_y_continuous(
    "Observations",
    expand = expansion(c(0.01, 0.1)),
    breaks = scales::extended_breaks(5)
  ) +
  scale_x_continuous(
    "Mood",
    expand = expansion(c(0.01))
  ) +
  facet_wrap("post", ncol = 1, scales = "free_y") +
  theme(legend.position = "none")

# Plot: Person-session-mean mood histograms at pre and during play
p_mood_prepost_sessions <- p_mood_prepost_raw %+%
  summarise(
    p_mood_prepost_raw$data,
    mood = mean(mood, na.rm = TRUE), .by = c(post, pid, ps)
  ) +
  scale_y_continuous(
    "Sessions",
    expand = expansion(c(0.01, 0.1)),
    breaks = scales::extended_breaks(5)
  )

# Plot: Person-mean mood histograms at pre and during play
p_mood_prepost_players <- p_mood_prepost_sessions %+%
  summarise(
    p_mood_prepost_raw$data,
    mood = mean(mood, na.rm = TRUE), .by = c(post, pid)
  ) +
  scale_y_continuous(
    "Players",
    expand = expansion(c(0.01, 0.1)),
    breaks = scales::extended_breaks(5)
  )

# Plot: Difference histogram (sessions)
p_mood_difference_sessions <- p_mood_prepost_sessions$data |>
  pivot_wider(names_from = post, values_from = mood) |>
  mutate(Difference = `During play` - `Pre-play`) |>
  drop_na(Difference) |>
  ggplot(aes(Difference)) +
  geom_histogram(bins = 50, col = "white") +
  geom_vline(xintercept = 0, linewidth = .5, col = "#2ca25f") +
  scale_y_continuous(
    "Sessions",
    expand = expansion(c(0.01, 0.1))
  ) +
  scale_x_continuous(
    expand = expansion(c(0.01))
  ) +
  coord_cartesian(xlim = c(-.4, .6))

# Plot: Difference histogram (players)
p_mood_difference_players <- p_mood_prepost_players$data |>
  pivot_wider(names_from = post, values_from = mood) |>
  mutate(Difference = `During play` - `Pre-play`) |>
  drop_na(Difference) |>
  ggplot(aes(Difference)) +
  geom_histogram(bins = 50, col = "white") +
  geom_vline(xintercept = 0, linewidth = .5, col = "#2ca25f") +
  scale_y_continuous(
    "Players",
    expand = expansion(c(0.01, 0.1))
  ) +
  scale_x_continuous(
    expand = expansion(c(0.01))
  ) +
  coord_cartesian(xlim = c(-.4, .6))

@@

```{r}
#| label: fig-ct1
#| include: true
#| fig.width: 4
#| fig.cap: "Estimated (changes) in mood as a function of session duration. Top: Average mood during a gaming session. Bottom: Change in mood during a session compared to mood at the beginning of a session. Gray ribbons indicate 95\\% confidence bands. We truncated the x-axis at three hours for this figure."

@@

```{r}
#| label: fig-ct2
#| fig.env: "figure*"
#| include: true
#| fig.asp: 0.5
#| fig.cap: "Estimated (changes) in mood as a function of session duration and pre-play mood. Top. Mood for the average player during a gaming session with pre-play mood at 5th, 25th, 50th, and 75th percentiles (columns). Ribbons indicate 95\\% confidence. Bottom. Same as above but with change in mood on the y-axis."
